On the probability of finding a nonphysical solution through shadowing

### 2020-10-26 main.tex
Edited most sections. Last pass before submitting.

\begin{figure}\centering
\includegraphics[width=0.48\textwidth]{lorenz_zmax_10_28_2.6666666666666665.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{lorenz_zmax2_10_28_2.6666666666666665.png}
\caption{L: the Lorenz map.  The x-axis is the $n$-th local maximum of $u_3(t)$ over a
long solution; the y-axis is the $(n+1)$-th local maximum of $u_3(t)$.  The intersection of this curve
		with the dashed line ($y=x$) indicates the initial condition for the solid line
in Figure \ref{fig:lorenz_periodic}.
R: the Lorenz map iterated twice.  The x-axis is the $n$-th local maximum
of $u_3(t)$; the y-axis is the $(n+2)$-th maximum.  The intersections with the diagonal dashed
line indicate the initial conditions for both the dotted and dashed lines in
Figure \ref{fig:lorenz_periodic}.}
\label{fig:lorenz_map}
\end{figure}
The Lorenz map provides us a tool to find as many periodic solutions as we want.
By intersecting the map with a diagonal line, we can find a local maximum of $u_3(t)$
for which the next maximum is almost the same value.  We then look up the values of
$u_1(t)$ and $u_2(t)$ when $u_3(t)$ achieved this maximum.  This gives us an initial
condition starting from which the solution is nearly periodic.   We can similarly
intersect the second iterate of the Lorenz map (right plot of Figure \ref{fig:lorenz_map},)
and the third iterate, etc, with a diagonal line to find increasingly
complex periodic solutions.

This Lorenz map is more than a tool to study the Lorenz equation.  It is a chaotic
dynamical system all by itself. Unlike the Lorenz equation, which is a continuous-time
dynamical system in three dimensions, the Lorenz map is a discrete-time dynamical
system in one dimension.  It exhibits the same sensitivity to initial condition
that characterizes the Lorenz equation.  One can readily observe in the right
plot of Figure \ref{fig:lorenz_map} that a small perturbation in the x-axis
can lead to a large change in the y-axis. This sensitivity grows exponentially
for the $n$-th iterate of this map. A solution to the Lorenz map can be obtained
by extracting the consecutive local maxima of a solution to the Lorenz equation.
If we extract from a physical solution to the Lorenz equation, we obtain a physical solution of the Lorenz map.  It will visit the
interval between 29 and 49 with varying, well-defined frequencies.
By contrast, if one extracts a solution to the Lorenz map from a periodic solution
to the Lorenz equation, the solution will visit only a discrete set of points.
It is thus a periodic, nonphysical solution.  What we learned about
the Lorenz equation could have all been learned from the Lorenz map.

Having discussed periodic solutions in this section,
we move to a second type of nonphysical solutions.
This type is more difficult to find and study than the periodic solutions.
To make it easier, we switch our example to a one-dimensional, discrete time dynamical
systems like the Lorenz map.  Because the Lorenz map lacks a closed form,
we construct, in the next subsection, a one-dimensional discrete-time dynamical system
with a closed form, one that qualitatively resembles the Lorenz map.  This map
will help us, in section \ref{sec:quasiphysical}, to study a more insidious type
of nonphysical solutions, a type that hides in the shadow.

\subsection{Tent map: periodic and physical solutions}
\label{sec:tent}
The tent map has been widely studied \cite{tent1}\cite{tent2}.  It is qualitatively similar
to the Lorenz map and has a simple, analytic form:
\begin{equation} \label{tentmap}
    \varphi(x) := \begin{cases}
    2x \quad & x < 1 \\
    4 - 2x \quad & 2 \ge x \ge 1
    \end{cases}
\end{equation}
Figure \ref{fig:tent_map} shows the tent map.
\begin{figure}\centering
\includegraphics[width=0.48\textwidth]{tent_map.png}
\caption{The tent map $\varphi$.}
\label{fig:tent_map}
\end{figure}

@@

It is easy to find periodic solutions for the tent map.
$\frac43$ maps to itself; $\frac45$ and $\frac85$ map to each other;
$\frac47, \frac87$, and $\frac{12}7$ map circularly.  In fact, any rational number evolves into periodic solutions that visit
only a finite set of rational numbers with the same denominator.
This is compatible
with ergodic theory because all rational numbers in $[0,2]$ comprise a subset of Lebesgue measure zero. That is, we would have zero likelihood of getting a
rational number if sampled the Lebesgue measure (the uniform distribution) on $[0,2]$.  Instead, with a 100\% probability, one would get an irrational number
that, when iterated under the tent map, leads to a physical trajectory that distributes uniformly on the interval $[0,2]$.

To understand why a physical solution visits the interval $[0,2]$ at a uniform
frequency, it is helpful to view the tent map from a different perspective.
For a randomly chosen $x_0\in[0,2]$, we can represent it in binary form:
\begin{equation}
    x_0 = \sum_{j=0}^{\infty} \frac{x_{0,j}}{2^j}
\end{equation}
where $x_{0,0}\in\{0,1\}$ is the integer component of $x_0$ and $x_{0,j}\in\{0,1\}$ is the $j$th bit after the binary point.
Let $x_{i+1} = \varphi(x_i)$, then it is straightforward to verify from the definition
of the map that its binary representation
\begin{equation}
    x_i = \sum_{j=0}^{\infty} \frac{x_{i,j}}{2^j}
\end{equation}
would satisfy
\begin{equation} \label{tent_xor}
    x_{i+1,j} = x_{i,0} \veebar x_{i,j+1}
\end{equation}
where $\veebar$ is the xor operator. To see why, note that 
multiplication by 2 is a left-shift operator in binary, and 
subtraction from 4 flips every bit after the binary point. If $x_0$ is chosen uniformly in $[0,2]$, then each bit $x_{0,j},j=0,1,\ldots$ of $x_0$
has equal probability of being 0 or 1, and each bit is independent of other bits.
It follows from Eq. \ref{tent_xor} that each bit $x_{i,j},j=0,1,\ldots$ of $x_i,
i=1,2,\ldots$ has equal probability of being 0 or 1,
and each bit is still independent of the other bits.
Thus, every $x_i$ in this physical solution is uniformly distributed in $[0,2]$.
As the map iterates starting from almost any $x_0$, a physical solution explores the entire interval $[0,2]$ uniformly.

\subsection{Nonphysical solutions Type II: Quasi-physical Solutions}
\label{sec:quasiphysical}
The simplicity of the tent map, as well as its binary form (\ref{tent_xor}),
enables us to study nonphysical solutions that are not periodic.
As in the last section, consider an
\begin{equation}
    x_0 = \sum_{j=0}^{\infty} \frac{x_{0,j}}{2^j}
\end{equation}
in which the bits $x_{0,j}$ are not independent of each other.  Instead,
suppose each bit is more likely to be identical to the previous bit than to be different.
That is, $x_{0,j+1}=x_{0,j}$ with probability $p>\frac12$ for $j=0,1,2,\ldots$.
Then, by Eq. \ref{tent_xor}, the bits of every $x_i, i=1,2,\ldots$ follow the same pattern, namely, each bit repeats the previous bit
with probability $p$. Moreover, consider Eq. \ref{tent_xor} 
for $j=0$ and any $i$:
\begin{equation}
    x_{i+1,0} = x_{i,0} \veebar x_{i,1}
\end{equation}
Because $x_{i,0}=x_{i,1}$ with probability $p>\frac12$, $x_{i+1,0}=0$ with probability 
$p>\frac12$.  Starting from $x_1$, this solution visits $[0,1]$ with
probability $p>\frac12$.  Instead of visiting $[0,1]$ and $[1,2]$ with
equal probability, as a physical solution does, this solution favors $[0,1]$. Since it provably visits the interval $[0,2]$ with a different frequency from that
of a physical solution, this is a nonphysical solution.

If $p = 1$, the bits of $x_0$ are either all zeros or all ones. So, all further iterates when $x_0$ is 0, are 0, and this is a trivial nonphysical solution of type 1. When $x_0$ consists of all ones, the solution eventually becomes 0, and hence is also a nonphysical solution that is eventually periodic (of period 1). Even when this solution does not attain zero in a large number of iterations, it does not 
distribute uniformly on $[0,2]$ like a physical solution 
because all $x_i, i\geq 1$ are of the form $1/2^k$ for some $k \in \mathbb{N}$; other numbers are not visited. Now when $p > \frac12$ but strictly less than 1, 
the solutions we just constructed are both nonphysical and aperiodic.
The bits of $x_0$, though correlated with each other,
still can exhibit an infinite variety of patterns. 
This implies that the solution may not eventually converge to 
any fixed point, nor show any repetitive patterns.  
 As the map iterates, these patterns
shift towards more significant digits, and the solution visits an infinite set of
points. 
 Nevertheless, the solutions observed along a trajectory also 
 do not conform to a uniform distribution on $[0,2]$ since they 
 preferentially visit the first half of this interval.
We call such aperiodic nonphysical solutions ``quasi-physical'' solutions.

\begin{figure}\centering
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.5001.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.51.png}
\\ \vspace{0.05\textwidth}
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.55.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.9.png}
\caption{Empirical density function of long solutions (a billion steps)
starting from four points whose binary digits have probability $p$ of
repeating the previous digit.
The solution shown in the top-left plot starts from a point with $p=0.5001$;
top-right: $p=0.51$; bottom-left: $p=0.55$; bottom-right: $p=0.9$.
Note that the empirical densities at larger $p'$s are well above 1.}
\label{fig:tent_quasiphysical}
\end{figure}

Figure \ref{fig:tent_quasiphysical} shows the empirical density functions of
four such quasi-physical solutions.  When $p=0.5001$,
the statistical distribution of the quasi-physical solution
is approximately uniform.  Recall that a physical solution explores $[0,2]$
uniformly.  In this case, the quasi-physical solution has very similar
statistical behavior as a physical solution.  When $p=0.51$, the empirical
distribution becomes ``hairy''.  An apparently fractal pattern emerges.
This fractal pattern further amplifies when $p=0.55$.  Meanwhile,
the density on the left side, $[0,1]$, becomes obviously higher than
the density on the right side, $[1,2]$.  This is consistent with our
theoretical analysis at the beginning of this subsection.
When $p=0.9$, the fractal pattern is so intensified that most
of the solution seem to concentrate in a collection of tiny intervals.
These plots expose the diversity of quasi-physical solutions.
Their statistical distribution can be either
very similar or completely different from that of physical solutions.

So far we have constructed one class of quasi-physical solutions.
It is noteworthy that there are infinite ways to 
construct quasi-physical solutions.  In the binary representation 
of the initial condition $x_0$, any statistical
deviation from equal probability of 0 or 1, or any statistical
dependence among digits would lead to quasi-physical solutions.
One could, for example, construct an $x_0$ in which a bit is
more likely to be 1 than 0 only if it follows two consecutive 0's.
Such $x_0$ would lead to a nonphysical solution.  Its statistical
distribution would differ from any of the plots in Figure
\ref{fig:tent_quasiphysical}.  Nevertheless, as can be
observed in Figure \ref{fig:tent_alternative}, it shows
a remarkable resemblance in its ``hairiness''; some kind of fractal
pattern emerges from the distribution.  A fractal distribution
appears to be a signature of quasi-physical solutions.
\begin{figure}\centering
    \centering
    \includegraphics[width=0.48\textwidth]{tent_quasiphysical_alternaive1_hist_p_0.6.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{tent_quasiphysical_alternaive2_hist_p_0.6.png}
    \caption{Empirical distribution function of other quasi-physical solutions.
    The solution shown in the left plot starts from an initial condition
    whose bits are independent and have probability 0.6 of being 0.
    The solution shown in the right plot starts from an initial condition
    whose bits have a probability of 0.6 of being 1 only following two
    consecutive 1's; otherwise a bit is 0 or 1 with equal probability.}
    \label{fig:tent_alternative}
\end{figure}

@@

Through a rigorous counter-example, we show that shadowing can lead
to nonphysical solutions. This has a troubling implication for numerical simulations of chaotic
governing equations.  Through the examples in this paper, it is clear that a numerical solution, satisfying
a governing equation perturbed due to numerics, is not 
expected to be shadowed by a
physical solution of the real governing physics.  What is, if any, the
relation of a numerical solution, such as DNS of turbulent flows, to
the true physics? For an ${\cal O}(10^{-18})$ 
perturbation (due to numerical error), 
we may get completely different statistics from a true, physical solution. How then can we trust numerical solutions when they are not guaranteed to share the long-term statistical behavior of the governing equation? 
We conclude with an example that illustrates a counterintuitive feature of some chaotic systems: a small perturbation to the 
governing equation can significantly change the statistics/long-term 
behavior of its physical solutions. We construct yet another 
perturbed tent map -- the \emph{plucked} 
tent map -- in which an oscillatory perturbation is introduced (see Supplementary Material section \ref{sec:pluckedTentMap} for the map equation). In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by a parameter $s$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map. The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.5$, are shown in Figure \ref{fig:pluckedTentMapDistribution}. On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$. On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe a fractal distribution.

In view of the plucked tent map, we can question further the
validity of numerical solutions. Can numerical play 
the role of the plucking perturbation to the 
governing equation? That is, can numerical solutions represent 
slight perturbations to the governing equations that nevertheless 
exhibit drastically ensemble statistics? Uniformly hyperbolic 
dynamical systems exhibit \emph{linear response} 
(\cite{ruelle}\cite{baladi}) by which small parameter perturbations 
lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value. But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely 
\cite{wormell}\cite{luca}. The plucked tent map provides a pathological example of a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).

Both the nonphysicality of shadowing solutions, and the existence of pathological statistical response, undermine the validity of using shadowing for sensitivity analysis of statistics \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. When the goal is to compute derivatives of ensemble averages, where the ensemble
is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results. This is because shadowing-based methods compute the sensitivities of ensemble statistics along shadowing solutions, but these may not be physical solutions that reproduce ensemble statistics. The error in shadowing sensitivities has been observed
before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. 
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{osc_tent_n.png}
    \includegraphics[width=0.45\textwidth]{osc_tent_s.png}
    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$). On the left, $s$ is held constant at 0.5. On the right, $n$ is held constant at 0.}
    \label{fig:pluckedTentMap}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_0.png}
    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_6.png}
    \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map. L: The stationary probability distribution of the plucked tent map at $s=0.5, n = 0$. R: The stationary probability distribution of the plucked tent map at $s=0.5, n = 6$.}
    \label{fig:pluckedTentMapDistribution}
\end{figure}
Violations of smooth response, like in the plucked tent map, has an immediate implication for shadowing sensitivities as well. In particular, if perturbed solutions are shadowed by physical solutions for all time, the change in the statistics of the perturbed solutions must be small. In other words, the physicality of shadowing predicts that there cannot be a large change in statistics due to small parameter perturbations; the reality is that, as illustrated by the plucked tent map, the effect of small parameter changes on the statistics can be drastic.

>>>>

\begin{figure}\centering
\includegraphics[width=0.48\textwidth]{lorenz_zmax_10_28_2.6666666666666665.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{lorenz_zmax2_10_28_2.6666666666666665.png}
\caption{L: the Lorenz map.  The x-axis is the $n$-th local maximum of $u_3(t)$ over a
long solution; the y-axis is the $(n+1)$-th local maximum of $u_3(t)$.  The intersection of this curve
		with the dashed line ($y=x$) indicates the initial condition for the solid line
in Figure \ref{fig:lorenz_periodic}.
R: the Lorenz map iterated twice.  The x-axis is the $n$-th local maximum
of $u_3(t)$; the y-axis is the $(n+2)$-th maximum.  The intersections with the diagonal dashed
line indicate the initial conditions for both the dotted and dashed lines in
Figure \ref{fig:lorenz_periodic}.}
\label{fig:lorenz_map}
\end{figure}
The Lorenz map provides us a tool to find as many periodic solutions as we want.
By intersecting the map with a diagonal line, we can find a local maximum of $u_3(t)$
for which the next maximum is almost the same value.  We then look up the values of
$u_1(t)$ and $u_2(t)$ when $u_3(t)$ achieves this maximum.  This gives us an initial
condition starting from which the solution is nearly periodic.   We can similarly
intersect the second iterate of the Lorenz map (right plot of Figure \ref{fig:lorenz_map}),
and the third iterate, etc, with a diagonal line, to find increasingly
complex periodic solutions.

This Lorenz map is more than a tool to study the Lorenz equation.  It is a chaotic
dynamical system all by itself. Unlike the Lorenz equation, which is a continuous-time
dynamical system in three dimensions, the Lorenz map is a discrete-time dynamical
system in one dimension.  It exhibits the same sensitivity to initial condition
that characterizes the Lorenz equation.  One can readily observe in the right
plot of Figure \ref{fig:lorenz_map} that a small perturbation in the x-axis
can lead to a large change in the y-axis. This sensitivity grows exponentially
for iterates of this map. A solution to the Lorenz map can be obtained
by extracting the consecutive local maxima of a solution to the Lorenz equation.
If we extract from a physical solution to the Lorenz equation, we obtain a physical solution of the Lorenz map.  It will visit the
interval between 29 and 49 with varying, well-defined frequencies.
By contrast, if one extracts a solution to the Lorenz map from a periodic solution
to the Lorenz equation, the solution will visit only a discrete set of points.
It is thus a periodic, nonphysical solution.  What we learned about
the Lorenz equation could have all been learned from the Lorenz map.

Having discussed periodic solutions in this section,
we move to a second type of nonphysical solutions.
This type is more difficult to find and study than the periodic solutions.
To make it easier, we switch our example to a one-dimensional, discrete time dynamical
systems like the Lorenz map.  Because the Lorenz map lacks a closed form,
we construct, in the next subsection, a one-dimensional discrete-time dynamical system
with a closed form, one that qualitatively resembles the Lorenz map.  This map
will help us, in section \ref{sec:quasiphysical}, to study a more insidious type
of nonphysical solutions that hides in shadowing solutions.

\subsection{Tent map: periodic and physical solutions}
\label{sec:tent}
The tent map has been widely studied \cite{tent1}\cite{tent2}.  It is qualitatively similar
to the Lorenz map and has a simple analytical form:
\begin{equation} \label{tentmap}
    \varphi(x) := \begin{cases}
    2x \quad & x < 1 \\
    4 - 2x \quad & 2 \ge x \ge 1
    \end{cases}
\end{equation}
Figure \ref{fig:tent_map} shows the tent map.
\begin{figure}\centering
\includegraphics[width=0.48\textwidth]{tent_map.png}
\caption{The tent map $\varphi$.}
\label{fig:tent_map}
\end{figure}

@@

It is easy to find periodic solutions for the tent map.
$\frac43$ maps to itself; $\frac45$ and $\frac85$ map to each other;
$\frac47, \frac87$, and $\frac{12}7$ map circularly.  In fact, any rational number evolves into periodic solutions that visit
only a finite set of rational numbers with the same denominator.
This is compatible
with ergodic theory because all rational numbers in $[0,2]$ comprise a subset of Lebesgue measure zero. That is, we would have zero likelihood of getting a
rational number if we sampled the Lebesgue measure (the uniform distribution) on $[0,2]$.  Instead, with a 100\% probability, one would get an irrational number
that, when iterated under the tent map, leads to a physical trajectory that distributes uniformly on the interval $[0,2]$.

To understand why a physical solution visits the interval $[0,2]$ at a uniform
frequency, it is helpful to view the tent map from a different perspective.
For a randomly chosen $x_0\in[0,2]$, we can represent it in binary form:
\begin{equation}
    x_0 = \sum_{j=0}^{\infty} \frac{x_{0,j}}{2^j}
\end{equation}
where $x_{0,0}\in\{0,1\}$ is the integer component of $x_0$ and $x_{0,j}\in\{0,1\}$ is the $j$th bit after the binary point.
Let $x_{i+1} = \varphi(x_i)$, then it is straightforward to verify from the definition
of the map that its binary representation
\begin{equation}
    x_i = \sum_{j=0}^{\infty} \frac{x_{i,j}}{2^j}
\end{equation}
satisfies
\begin{equation} \label{tent_xor}
    x_{i+1,j} = x_{i,0} \veebar x_{i,j+1}
\end{equation}
where $\veebar$ is the xor operator. To see why, note that 
multiplication by 2 is a left-shift operator in binary, and 
subtraction from 4 flips every bit after the binary point. If $x_0$ is chosen uniformly in $[0,2]$, then each bit $x_{0,j},j=0,1,\ldots$ of $x_0$
has equal probability of being 0 or 1, and each bit is independent of other bits.
It follows from Eq. \ref{tent_xor} that each bit $x_{i,j},j=0,1,\ldots$ of $x_i,
i=1,2,\ldots$ has equal probability of being 0 or 1,
and each bit is still independent of the other bits.
As the map iterates starting from almost any $x_0$, a physical solution explores the entire interval $[0,2]$ uniformly.

\subsection{Nonphysical solutions Type II: Quasi-physical Solutions}
\label{sec:quasiphysical}
The simplicity of the tent map, as well as its binary form (Eq. \ref{tent_xor}),
enables us to study nonphysical solutions that are not periodic.
As in the last section, consider an
\begin{equation}
    x_0 = \sum_{j=0}^{\infty} \frac{x_{0,j}}{2^j}
\end{equation}
in which the bits $x_{0,j}$ are not independent of each other.  Instead,
suppose each bit is more likely to be identical to the previous bit than to be different.
That is, $x_{0,j+1}=x_{0,j}$ with probability $p>\frac12$ for $j=0,1,2,\ldots$.
Then, by Eq. \ref{tent_xor}, the bits of every $x_i, i=1,2,\ldots$ follow the same pattern, namely, each bit repeats the previous bit
with probability $p$. Moreover, consider Eq. \ref{tent_xor} 
for $j=0$ and any $i$:
\begin{equation}
    x_{i+1,0} = x_{i,0} \veebar x_{i,1}.
\end{equation}
Because $x_{i,0}=x_{i,1}$ with probability $p>\frac12$, $x_{i+1,0}=0$ with probability 
$p>\frac12$.  Starting from $x_1$, this solution visits $[0,1]$ with
probability $p>\frac12$.  Instead of visiting $[0,1]$ and $[1,2]$ with
equal probability, as a physical solution does, this solution favors $[0,1]$. Since it provably visits the interval $[0,2]$ with a different frequency from that
of a physical solution, this is a nonphysical solution.

If $p = 1$, the bits of $x_0$ are either all zeros or all ones, which is the same as all zeros in mod 2 arithmetic. So, all further iterates when $x_0$ is 0, are 0, and this is a trivial nonphysical solution of type 1.  Now when $p > \frac12$ but strictly less than 1, 
the solutions we just constructed are both nonphysical and aperiodic.
The bits of $x_0$, though correlated with each other,
still can exhibit an infinite variety of patterns. 
This implies that the solution may not eventually converge to 
any fixed point, nor show any repetitive patterns.  
 As the map iterates, these patterns
shift towards more significant digits, and the solution visits an infinite set of
points. 
 Nevertheless, the solutions observed along a trajectory also 
 do not conform to a uniform distribution on $[0,2]$ since they 
 preferentially visit the first half of this interval.
We call such aperiodic nonphysical solutions ``quasi-physical'' solutions.

\begin{figure}\centering
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.5001.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.51.png}
\\ \vspace{0.05\textwidth}
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.55.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.9.png}
\caption{Empirical probability distribution of long solutions (a billion steps)
starting from four points whose binary digits have probability $p$ of
repeating the previous digit.
The solution shown in the top-left plot starts from a point with $p=0.5001$;
top-right: $p=0.51$; bottom-left: $p=0.55$; bottom-right: $p=0.9$.
}
\label{fig:tent_quasiphysical}
\end{figure}

Figure \ref{fig:tent_quasiphysical} shows the empirical density functions of
four such quasi-physical solutions.  When $p=0.5001$,
the statistical distribution of the quasi-physical solution
is approximately uniform.  Recall that a physical solution explores $[0,2]$
uniformly.  In this case, the quasi-physical solution has very similar
statistical behavior as a physical solution.  When $p=0.51$, the empirical
distribution becomes ``hairy''.  An apparently fractal pattern emerges.
This fractal pattern further amplifies when $p=0.55$.  Meanwhile,
the density on the left side, $[0,1]$, becomes obviously higher than
the density on the right side, $[1,2]$.  This is consistent with our
theoretical analysis at the beginning of this subsection.
When $p=0.9$, the fractal pattern is so intensified that most
of the solution seems to concentrate in a collection of tiny intervals.
These plots expose the diversity of quasi-physical solutions.
Their statistical distribution can be either
very similar or completely different from that of physical solutions.

So far we have constructed one class of quasi-physical solutions.
It is noteworthy that there are infinite ways to 
construct quasi-physical solutions.  In the binary representation 
of the initial condition $x_0$, any statistical
deviation from equal probability of 0 or 1, or any statistical
dependence among digits would lead to quasi-physical solutions.
One could, for example, construct an $x_0$ in which a bit is
more likely to be 1 than 0 only if it follows two consecutive 0's.
Such an $x_0$ would lead to a nonphysical solution.  Its statistical
distribution would differ from any of the plots in Figure
\ref{fig:tent_quasiphysical}.  Nevertheless, as can be
observed in Figure \ref{fig:tent_alternative}, it shows
a remarkable resemblance in its ``hairiness''; some kind of fractal
pattern emerges from the distribution.  A fractal distribution
appears to be a signature of quasi-physical solutions.
\begin{figure}\centering
    \centering
    \includegraphics[width=0.48\textwidth]{tent_quasiphysical_alternaive1_hist_p_0.6.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{tent_quasiphysical_alternaive2_hist_p_0.6.png}
    \caption{Empirical distribution function of other quasi-physical solutions.
    The solution shown in the left plot starts from an initial condition
    whose bits are independent and have probability 0.6 of being 0.
    The solution shown in the right plot starts from an initial condition
    whose bits have a probability of 0.6 of being 1 only following two
    consecutive 1's; otherwise a bit is 0 or 1 with equal probability.}
    \label{fig:tent_alternative}
\end{figure}

@@

Through a rigorous counter-example, we show that shadowing can lead
to nonphysical solutions. This has a troubling implication for numerical simulations of chaotic
governing equations.  Through the examples in this paper, it is clear that a numerical solution, satisfying
a governing equation perturbed due to numerics, is not 
expected to be shadowed by a
physical solution of the real governing physics.  What is, if any, the
relation of a numerical solution, such as DNS of turbulent flows, to
the true physics? For an ${\cal O}(10^{-18})$ 
perturbation (due to numerical error), 
we may get completely different statistics from a true, physical solution. How then can we trust numerical solutions when they are not guaranteed to share the long-term statistical behavior of the governing equation? 
We conclude with an example that illustrates a counterintuitive feature of some chaotic systems: a small perturbation to the 
governing equation can significantly change the statistics/long-term 
behavior of its physical solutions. We construct yet another 
perturbed tent map -- the \emph{plucked} 
tent map -- in which an oscillatory perturbation is introduced (see Supplementary Material section \ref{sec:pluckedTentMap} for the map equation). In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by a parameter $s$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map. The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.5$, are shown in Figure \ref{fig:pluckedTentMapDistribution}. On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$. On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparently fractal distribution reminiscent of that of the quasi-physical solutions in 
section \ref{sec:tentPerts}.

In view of the plucked tent map, we can question further the
validity of numerical solutions. Can numerical play 
the role of the plucking perturbation to the 
governing equation? That is, can numerical solutions represent 
slight perturbations to the governing equations that nevertheless 
exhibit drastically different statistics from that of the ensemble? Uniformly hyperbolic 
dynamical systems exhibit \emph{linear response} 
(\cite{ruelle}\cite{baladi}) by which small parameter perturbations 
lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value. But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely 
\cite{wormell}\cite{luca}. The plucked tent map provides a pathological example of a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).

Both the nonphysicality of shadowing solutions, and the existence of extremely non-smooth statistical response, undermine the validity of using shadowing for sensitivity analysis of statistics \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. When the goal is to compute derivatives of ensemble averages, where the ensemble
is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results. This is because shadowing-based methods compute the sensitivities of ensemble statistics along shadowing solutions, but these may not be physical solutions that reproduce ensemble statistics. The error in shadowing sensitivities has been observed
before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. 
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{osc_tent_n.png}
    \includegraphics[width=0.45\textwidth]{osc_tent_s.png}
    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$). On the left, $s$ is held constant at 0.5. On the right, $n$ is held constant at 0.}
    \label{fig:pluckedTentMap}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_0.png}
    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_6.png}
    \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map. L: The stationary probability distribution of the plucked tent map at $s=0.5, n = 0$. R: The stationary probability distribution of the plucked tent map at $s=0.5, n = 6$.}
    \label{fig:pluckedTentMapDistribution}
\end{figure}
Violations of smooth response, like in the plucked tent map, have an immediate implication for shadowing sensitivities as well. In particular, if perturbed solutions are shadowed by physical solutions for all time, the change in the statistics of the perturbed solutions must be small. In other words, the physicality of shadowing predicts that there cannot be a large change in statistics due to small parameter perturbations; the reality is that, as illustrated by the plucked tent map, the effect of small parameter changes on the statistics can be drastic.
